### 2017-10-06 Project1/LaTeX/report.tex
Added report

\documentclass[times]{article}

\usepackage[margin=1.0in]{geometry}
\usepackage{graphicx}
\usepackage{adjustbox}
\usepackage{float}
\usepackage{placeins}
\usepackage[none]{hyphenat}
\usepackage{amsmath}
\usepackage[us]{datetime}
\usepackage[explicit]{titlesec}
%\titleformat{\section}{\normalfont}{}{0em}{\textbf{\large Problem  \thesection}\  \normalsize #1}
\begin{document}
	\title{CS 6601 Secure Data Analysis - Fall 2017 \\ Project 1 \\ Millionaire Problem using Garbled Circuits}
	\author{Adam Bowers \\ Sammie Bush \\ Dalton Cole}
	\date{\formatdate{6}{10}{2017}}
	\maketitle

	\section{Implementation}
	Our main worker for this project is the \textit{h1Circuit.py} script. What this script does is create the input files for GCParser. The needed input files are: \textit{millionaireCircuit.cir}, \textit{partyOneInputFile}, and \textit{partyTwoInputFile}. The purpose of these files are: the main input file, the client's input, and the server's input, respectively.

	All results of the experiment can be seen in the result's directory. The \textit{siclientout} and \textit{siserverout} correspond party 1 and party 2, respectively.

\end{document}