### 2017-12-04 hw2/main.tex
Add files via upload

% !TEX TS-program = pdflatex
% !TEX encoding = UTF-8 Unicode

% This is a simple template for a LaTeX document using the "article" class.
% See "book", "report", "letter" for other types of document.

\documentclass[10pt]{article} % use larger type; default would be 10pt

\usepackage[utf8]{inputenc} % set input encoding (not needed with XeLaTeX)

%%% Examples of Article customizations
% These packages are optional, depending whether you want the features they provide.
% See the LaTeX Companion or other references for full information.

%%% PAGE DIMENSIONS
\usepackage{geometry} % to change the page dimensions
\geometry{a4paper} % or letterpaper (US) or a5paper or....
% \geometry{margin=2in} % for example, change the margins to 2 inches all round
% \geometry{landscape} % set up the page for landscape
%   read geometry.pdf for detailed page layout information

\usepackage{graphicx} % support the \includegraphics command and options

% \usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent

%%% PACKAGES
\usepackage{booktabs} % for much better looking tables
\usepackage{array} % for better arrays (eg matrices) in maths
\usepackage{paralist} % very flexible & customisable lists (eg. enumerate/itemize, etc.)
\usepackage{verbatim} % adds environment for commenting out blocks of text & for better verbatim
\usepackage{subfig} % make it possible to include more than one captioned figure/table in a single float
\usepackage{hyperref}
% These packages are all incorporated in the memoir class to one degree or another...

%%% HEADERS & FOOTERS
\usepackage{fancyhdr} % This should be set AFTER setting up the page geometry
\pagestyle{fancy} % options: empty , plain , fancy
\renewcommand{\headrulewidth}{0pt} % customise the layout...
\lhead{}\chead{}\rhead{}
\lfoot{}\cfoot{\thepage}\rfoot{}

%%% SECTION TITLE APPEARANCE
\usepackage{sectsty}
\allsectionsfont{\sffamily\mdseries\upshape} % (See the fntguide.pdf for font help)
% (This matches ConTeXt defaults)

%%% ToC (table of contents) APPEARANCE
\usepackage[nottoc,notlof,notlot]{tocbibind} % Put the bibliography in the ToC
\usepackage[titles,subfigure]{tocloft} % Alter the style of the Table of Contents
\renewcommand{\cftsecfont}{\rmfamily\mdseries\upshape}
\renewcommand{\cftsecpagefont}{\rmfamily\mdseries\upshape} % No bold!

%%% END Article customizations

%%% The "real" document content comes below...

\title{HW2}
\author{Adam Bowers, Dalton Cole, Sammie Bush}
%\date{} % Activate to display a given date or no date (if empty),
         % otherwise the current date is printed

\begin{document}
\maketitle

\subsection*{\textbf{PseudoCode}}
Let l be the bit length of u\&v, and let 1 be most significant bit and l least significant bit.
Part 1\\
Party one randomly chooses functionality F=u>v or u<v
He first computes 
\begin{itemize}
	\item for i=1 to l 
		\begin{itemize}
			\item $E_{pk} (u_i * v_i ) \gets SM(E_{pk} (u_i) , E_{pk}(v_i))$
			\\ first party one computes the product of current bit of v and u
			\item if selected F: was u$>$v
			 \subitem $W_i \gets E_{pk} (u_i) * E_{pk} (u_i * v_i)^{n-1}$
			 \\compute $u_i - u_i*v_i $ if there is any $W_i=1$ we know for i $u_i=1$ \& $v_i=0$
			 \subitem $\Gamma_i \gets E_{pk}(v_i - u_i) * E_{pk}(r_i)$ where $r_i \in Z_n$
			 \\compute $v_i - u_i +r_i$
			\item else
			\subitem $W_i \gets E_{pk} (v_i) * E_{pk} (u_i * v_i)^{N-1}$
			\\compute $v_i - u_i*v_i $  if there is any $W_i=1$ we know for i $v_i=1$ \& $u_i=0$
			\subitem $\Gamma_i \gets E_{pk}(u_i - v_i) * E_{pk}(r_i)$ where $r_i \in Z_n$
			\\compute $u_i - v_i +r_i$
			\item $G_i \gets E_{pk}(u_i \oplus v_i)$
			\\now compute xor of two bits if 0 they're the same so first 1 tells us first different bit
			\item $H_i \gets H_{i-1}^{s_i} * G_i ;$ where $s_i \in Z_n$ and $H_0 = E_{pk}(0)$
			\\now mask the xor based on previous bits and a random number, Hs will be 0 till first 1 then every term past that will be a random value based on $s$ values.
			\item $\Phi_i \gets E_{pk}(-1) * H_i$
			\\shift the domain of by n-1
			 \textbf{Not sure what this is for? to mask output for party two? two make sure 1s are 0 and every other value is a random value?}
			\item $L_i \gets W_i * \Phi_i^{t_i}$ where $t_i \in Z_n$
			\\ if $W_i$ is 1 $\Phi_i^{t_i}$ will be 0 thus $L_i$ will be 1 otherwise it will be some random value 
		\end{itemize}
	\item  $\Gamma' \gets \pi_1(\Gamma)$
	\item $L' \gets \pi_2(L)$
	\\we permute the outputs so P2 can't tell anything from indices of bits
	\item send $\Gamma' \& L'$ to P2
\end{itemize}
Part 2\\
Party two computes
\begin{itemize}
	\item $M \gets D_{sk}(L_i')$
	\\decrypt our Ls
	\item if there exists an $M_i=1$
		\subitem $\alpha \gets 1$
	\item else
		\subitem $\alpha \gets 0$
	\\assign alpha based on L values. If there is a 1 we know the selected F function of P1
	\\is true but not what it is
	\item $M_i' \gets \Gamma_i^{\alpha '} for 1\le i \le l$
	\\if we have a 1 in M then P1 gets $Gamma$ back, otherwise they get random values
	\item send $M'$ and $E_{pk}(\alpha)$ to P1
\end{itemize}
Part 3\\
\begin{itemize}
	\item $M \gets \pi^{-1}(M')$
	\\un permute M
	\item for i= 1 to l
		\subitem $\lambda_i \gets M_i * E_{pk}(\alpha)^{N-r_i}$
		\\if alpha = 1 we subtract the random value added to L in part one
		\\to get either $v-u$ or $u-v$ depending on f chosen
		\subitem if F = $u>v$ 
			\subitem $E_{pk}(min(u,v)_i) \gets E_{pk}(u_i) * \lambda_i$
			\\if $u>v$ $\lambda$ = $v-u$ if $v>u$ otherwise 0
			\\therefore $u_i +\lambda_i$ gives us bit of highest value
		\subitem else
			\subitem $E_{pk}(min(u,v)_i) \gets E_{pk}(v_i) * \lambda_i$
			\\if $v>u$ $\lambda$ = $u-v$ if $u>v$ otherwise 0
			\\therefore $v_i +\lambda_i$ gives us bit of highest value
	\item concat $E_{pk}(min(u,v)_i)$ and Party one has $E_{pk}(min(u,v))$ as required

\end{itemize}

\subsection*{\textbf{Two Phase Version}}
Let l be the bit length of u\&v, and let 1 be most significant bit and l least significant bit.
Furthermore let BD be a secure Binary Decomposition function \& SM be a secure multiplication function

Phase one of the protocol returns E(0) if u is the minimum and E(1) if v is the minimum or equal to u.

\textbf{(Minimum(E(u),E(v))}
\\where E(u) and E(v) are encrypted bitwise
Part 1\\
Party one randomly chooses functionality F=u>v or u<v
He first computes 
\begin{itemize}
	\item for i=1 to l 
	\begin{itemize}
		\item $E_{pk} (u_i * v_i ) \gets SM(E_{pk} (u_i) , E_{pk}(v_i))$
		\\ first party one computes the product of current bit of v and u
		\item if selected F: was u$>$v
		\subitem $W_i \gets E_{pk} (u_i) * E_{pk} (u_i * v_i)^{n-1}$
		\\compute $u_i - u_i*v_i $ if there is any $W_i=1$ we know for i $u_i=1$ \& $v_i=0$
		\item else
		\subitem $W_i \gets E_{pk} (v_i) * E_{pk} (u_i * v_i)^{N-1}$
		\\compute $v_i - u_i*v_i $  if there is any $W_i=1$ we know for i $v_i=1$ \& $u_i=0$
		\item $G_i \gets E_{pk}(u_i \oplus v_i)$
		\\now compute xor of two bits if 0 they're the same so first 1 tells us first different bit
		\item $H_i \gets H_{i-1}^{s_i} * G_i ;$ where $s_i \in Z_n$ and $H_0 = E_{pk}(0)$
		\\now mask the xor based on previous bits and a random number, Hs will be 0 till first 1 then every term past that will be a random value based on $s$ values.
		\item $\Phi_i \gets E_{pk}(-1) * H_i$
		\\shift the domain of by n-1
		\textbf{Not sure what this is for? to mask output for party two? two make sure 1s are 0 and every other value is a random value?}
		\item $L_i \gets W_i * \Phi_i^{t_i}$ where $t_i \in Z_n$
		\\ if $W_i$ is 1 $\Phi_i^{t_i}$ will be 0 thus $L_i$ will be 1 otherwise it will be some random value 
	\end{itemize}
	\item $L' \gets \pi_2(L)$
	\\we permute the output so P2 can't tell anything from indices of bits
	\item send $L'$ to P2
\end{itemize}
Part 2\\
Party two computes
\begin{itemize}
	\item $M \gets D_{sk}(L_i')$
	\\decrypt our Ls
	\item if there exists an $M_i=1$
	\subitem $\alpha \gets 1$
	\item else
	\subitem $\alpha \gets 0$
	\\assign alpha based on L values. If there is a 1 we know the selected F function of P1
	\\is true but not what it is
	\item send $E_{pk}(\alpha)$ to P1
\end{itemize}
Part 3\\
\\if alpha is one function is true 0 false, map that to
\\0 if u is min otherwise 1
\begin{itemize}
	\item if F = $u>v$  
		\subitem $E_{pk}(result) \gets E_{pk}(alpha)$
	\item else	
		\subitem $E_{pk}(result) \gets (E_{pk}(alpha)*E_pk(-1))^{N-1}$
	\\if F was $u>v$ alpha already has 0 if u is min 1 if v
	\\for the other function we have to subtract 1 and to get 0 if u is minimum 1 otherwise
	\\return $E_pk(result)$
\end{itemize}

In phase two party one uses minimum function to get min index and returns that value
\textbf{Return Minimum}\\
P1 computes
\begin{itemize}
	\item $E_{pk}(u)  = BD(E_pk(u))$
	\item $E_{pk}(v)  = BD(E_pk(v))$
	\item $MinIndice = Minimum(E_{pk}(u),E_{pk}(v))$
	\item for i=1 to l
 		\subitem $\Gamma_i \gets E_{pk}(v_i - u_i) * E_{pk}(r_i)$ where $r_i \in Z_n$
 		\\compute $v_i - u_i +r_i$
 	\item  $\Gamma' \gets \pi_1(\Gamma)$
 	\item send $\Gamma'$ and $MinIndice$ to P2
\end{itemize}

P2 computes

\begin{itemize}
	\item $MinIndice \gets D_{sk}(MinIndice)$
	\item $M_i' \gets \Gamma_i^{(MinIndice}$ for $1\le i \le l$
	\item return $M'$
\end{itemize}

P1 computes

\begin{itemize}
\item $M \gets \pi^{-1}(M')$
\item $\lambda _i \gets M_i * E_{pk}(MinIndice)^{N-r_i}$ for i= 1 to l
\\remove random terms from the difference so we can return the minimum
\item $\Gamma \gets E_{pk}(u)*\lambda$
\\this returns u if u is minimum otherwise v
\item $E_{pk}(min(E_{pk}(u),E_{pk}(v))) = \Gamma$
\end{itemize}

Thus party one has minimum encrypted element between u \& v as required
\subsection*{\textbf{Description}}
Pre-work
Set a bit length for u and v.\\
Phase 1:\\
	Goal:
    To compute whether u or v is the minimum\\

	Part 1:
	First,Party 1 chooses a function. Either greater than or less than.
	Next, for each bit, we compute a list of masks L.
	Next, we permute these masks to create L'.
    Party 1 sends L' to Party 2.\\

	Part 2:
	Party 2 decrypts L' to create M.
    If any of the elements in M equals 1, $\alpha$, else it equals 0. If 1 is present, we know the function chosen in the 		beginning is true.
    Party 2 then sends an encrypted copy of $\alpha$ \\

	Part 3:
	If alpha is 1, function is true. If alpha is 0, it is false. Map that to 0 if u is min, otherwise, map to 1.
    If F was u > v alpha already has 0 if u is min 1 if v
for the other function we have to subtract 1 and to get 0 if u is minimum 1, otherwise, return Epk(result)\\

Phase 2:\\
	Goal:
    To return the calculated minimum to the requesting party.\\

    First, Party 1 computes Epk(u), Epk(v), and MinIndice.
    Next, for each bit, Party 1 computes $\Gamma$ and $\Gamma'$.
    Lastly, Party 1 sends $\Gamma'$ and MinIndice to Party 2.\\

	Party 2 then decrypts the received MinIndice.
    With that and $\Gamma$, they compute $M_i'$.
    Party 2 then sends M' to Party 1.\\

    Party 1 then computes M from M'.
    With that, Party 1 is able to compute $\lambda$ and $\Gamma$.
    For each $\lambda$, Party 2 removes random terms from the differences so that the minimum can be found.
    If $\Gamma$ is returned when computing it, u is the minimum. otherwise, v is the minimum.\\

\subsection*{\textbf{Example}}
The above algorithm is implemented at: \url{https://github.com/drc14/Secure_Minimum}. The example was generated using the program implemented.

\textbf{Minimum(E(u), E(v))} \\
\textbf{Party 1} \\
Choose random functionality f: u $>$ v \\
For 1 in 3: \\
\indent $E(u_1) = E(1)$ \\
\indent $E(v_1) = E(1)$ \\
\indent $E(u_1 * v_1) =  Secure_Multiplication(E(u_1), E(v_1)) = E(1)$ \\
\indent F: u $>$ v: \\
\indent \indent $W_1 = E(0)$ \\
\indent $G_1 = E(u_1 \oplus v_1) = E(0)$ \\
\indent $H_1 = E(0)$ \\
\indent $\Phi_1 = E(12)$ \\
\indent $L_1 = W_1 * \Phi^(t_1) = E(0) * E(12)^8 = E(5)$ \\
For 2 in 3: \\
\indent $E(u_2) = E(1)$ \\
\indent $E(v_2) = E(0)$ \\
\indent $E(u_2 * v_2) = Secure_Multiplication(E(u_2), E(v_2)) = E(0)$ \\
\indent F: u $>$ v: \\
\indent \indent $W_2 = E(1)$ \\
\indent $G_2 = E(u_2 \oplus v_2) = E(1)$ \\
\indent $H_2 = E(1)$ \\
\indent $\Phi_2 = E(0)$ \\
\indent $L_2 = W_2 * \Phi^(t_2) = E(1) * E(0)^{10} = E(1)$ \\
For 3 in 3: \\
\indent $E(u_3) = E(1)$ \\
\indent $E(v_3) = E(1)$ \\
\indent $E(u_3 * v_3) = Secure_Multiplication(E(u_3), E(v_3)) = E(1)$ \\
\indent F: u $>$ v: \\
\indent \indent $W_3 = E(0)$ \\
\indent $G_3 = E(u_3 \oplus v_3) = E(0)$ \\
\indent $H_3 = E(2)$ \\
\indent $\Phi_3 = E(1)$ \\
\indent $L_3 = W_3 * \Phi^(t_3) = E(0) * E(1)^{2} = E(2)$ \\
$E(L) = E(5, 1, 2)$ \\
$E(L') = E(2, 1, 5)$ \\ \\

\textbf{Party 2} \\
$M = D(E(L')) = [2, 1, 5]$ \\
$\alpha = 1$ because 1 appears in $M$\\
$E(\alpha) = E(1)$ \\ \\
\textbf{Party 1} \\
$F: u > v$ \\
\indent Return $E(\alpha) = E(1)$ \\ \\

\textbf{Party 1} \\
E(u) = BD(E(7)) \\ 
E(v) = BD(E(5)) \\
Minimum\_Index $= Mininimum(E(u), E(v)) = E(1)$ \\
For $i=1$ to 3:
\indent $Gamma_1 = E(v_1 - u_1) * E(r_1) = E(1 - 1) * E(5) = E(5)$ \\
\indent $Gamma_2 = E(v_2 - u_2) * E(r_2) = E(0 - 1) * E(10) = E(9)$ \\
\indent $Gamma_3 = E(v_3 - u_3) * E(r_3) = E(1 - 1) * E(8) = E(8)$ \\
$\Gamma = E(5, 9, 8)$ \\
$\Gamma' = \pi_1(\Gamma) = E(9, 8, 5)$ \\ \\

\textbf{Party 2} \\
\textit{$Min_Index = D(E(Min_Index)) = 1$} \\ 
$M' = E(9, 8, 5)^1 = E(9, 8, 5)$ \\ \\

\textbf{Party 1} \\
$M = \pi^{-1}(M') = E(5, 9, 8)$ \\
$\lambda = M * E(\textit{Min\_Index})^{N - r_i} = E(4, 11, 10)$ \\
$\Gamma = E(u) * \lambda = E(5)$ \\
$E(min(E(u), E(v))) = \Gamma = E(5)$ \\

\end{document}